### 2019-07-03 methodology/instruments.tex
Integrati TODO di AA per Cap. 3 e sezione su correlazione e IG

\section{Methods} 
\subsection{Libraries}
\subsubsection{pomegranate}
\textit{pomegranate} (\cite{pomegranate}) is an open-source probabilistic models package for python.
Its core philosophy is that every probabilistic model, from Hidden Markov to Bayesian Network, can be seen as a probability distribution and, as such, can be flexibly composed into hierarchical mixture models (\cite{Schreiber2017}).
The package implements:
\begin{itemize}
	\item Probability Distributions
	\item General Mixture Models
	\item Hidden Markov Models
	\item Bayes Classifiers and Na{\"i}ve Bayes
	\item Markov Chains
	\item \textbf{Bayesian Networks}
	\item Factor Graphs
\end{itemize}

@@

\subsubsection{Gurobi}
\textit{Gurobi} (\cite{gurobi}) is a closed-source mathematical programming solver for Linear Programming, Quadratic Programming and Mixed-Integer Programming optimisation problems.
It calims to be the fastest solver available.
Gurobi offers object-oriented and matrix-oriented interfaces to, among others, Python, MATLAB and Excel.

\subsubsection{pandas}
\textit{pandas} (\cite{pandas}) is an extremely widely-used open-source python library that provides data structures and methods to aid in data analysis.
The package excels in the manipulation of tabular data in the form of \texttt{DataFrame}, that is the analogous of R's \texttt{data.frame}.
A \texttt{DataFrame} can be seen as a \enquote{general 2D, size-mutable structure with potentially heterogeneously-typed columns}.
The syntax for slicing is very close to R's as are many other functionalities; this is because one of Pandas' explicit goals is to offer all of CRAN's functionalities and be easily approachable by anyone already knowing the other language.

@@

\subsubsection{scikit-learn}
\textit{scikit-learn} (\cite{scikitlearn}) aims at providing a unified API for basic Machine Learning; it does not include advanced paradigms such as Reinforcement Learning or graphical models for structured learning.
The latter omission was the reason that lead me to select pomegranate as the basis for the implementation of a Bayesian Network.
What is included are a stack supervised and unsupervised ML tools to prepare data sets, define machine learning models ranging from spectral analysis-based to ensemble methods to clustering and multiple evaluation and model selection utilities.

@@

\subsubsection{networkx}
\textit{NetworkX} (\cite{networkx}) is another widely-used package that is specialised in the creation and manipulation of graph-structured data.

\subsubsection{anytree}
\textit{anytree} (\cite{anytree}) is a python package giving a lightweight implementation of a tree data structure and traversal methods.

@@

The full procedure can be found in \cite{koller2007dseparation}; my implementation follows this pseudocode very closely but instead of finding all nodes $R$ that are d-connected to the input $X$, tests if a given target $Y$ is d-separated from $X$ or not.
This gives some extra flexibility in how the function can be used.
To find the set $S$ of all nodes d-separated from $X$ I simply iterate the test over all nodes in the graph.

\lstset{language=algebra,linewidth=0.95\linewidth,breaklines=true,numbers=left,
basicstyle=\ttfamily,numberstyle=\tiny,escapeinside={//*}{\^^M},
mathescape=true}
\begin{lstlisting}
source // source variable
evidence // evidence set
foreach target in $V$ // $V$ set of variables of the graph
	separated_from_source[target] = d_separated(source,target) // true or false for each variable
\end{lstlisting}

>>>>

\section{Methods} 
\subsection{Libraries}
\subsubsection{Pomegranate}
\textit{pomegranate} (\cite{pomegranate}) is an open-source probabilistic models package for python.
Its core philosophy is that every probabilistic model, from Hidden Markov to Bayesian Network, can be seen as a probability distribution and, as such, can be flexibly composed into hierarchical mixture models (\cite{Schreiber2017}).
The package implements:
\begin{itemize}
	\item Probability Distributions
	\item General Mixture Models
	\item Hidden Markov Models
	\item Bayes Classifiers and Na{\"i}ve Bayes
	\item Markov Chains
	\item \textbf{Bayesian Networks}
	\item Factor Graphs
\end{itemize}

@@

\subsubsection{Gurobi}
\textit{Gurobi} (\cite{gurobi}) is a closed-source mathematical programming solver for Linear Programming, Quadratic Programming and Mixed-Integer Programming optimisation problems.
It claims to be the fastest solver available for these classes of problems.
Gurobi offers object-oriented and matrix-oriented interfaces to, among others, Python, MATLAB and Excel.

\subsubsection{Pandas}
\textit{pandas} (\cite{pandas}) is an extremely widely-used open-source python library that provides data structures and methods to aid in data analysis.
The package excels in the manipulation of tabular data in the form of \texttt{DataFrame}, that is the analogous of R's \texttt{data.frame}.
A \texttt{DataFrame} can be seen as a \enquote{general 2D, size-mutable structure with potentially heterogeneously-typed columns}.
The syntax for slicing is very close to R's as are many other functionalities; this is because one of Pandas' explicit goals is to offer all of CRAN's functionalities and be easily approachable by anyone already knowing the other language.

@@

\subsubsection{Scikit-learn}
\textit{scikit-learn} (\cite{scikitlearn}) aims at providing a unified API for basic Machine Learning; it does not include advanced paradigms such as Reinforcement Learning or graphical models for structured learning.
The latter omission was the reason that lead me to select pomegranate as the basis for the implementation of a Bayesian Network.
What is included are a stack supervised and unsupervised ML tools to prepare data sets, define machine learning models ranging from spectral analysis-based to ensemble methods to clustering and multiple evaluation and model selection utilities.

@@

\subsubsection{Networkx}
\textit{NetworkX} (\cite{networkx}) is another widely-used package that is specialised in the creation and manipulation of graph-structured data.
The main use for this package was in building the \texttt{MPEGraph} ( \todo[inline]{inserire referenza una volta scritto} ) data structure that I used to build a graph representing the outcome of pseudo-MPE (see \ref{subsubsec:pseudo-mpe} ) applied to multiple initial evidences (see \ref{subsubsec: pseudo-mpe-random} ) .

\subsubsection{Anytree}
\textit{anytree} (\cite{anytree}) is a python package providing a lightweight implementation of a tree data structure and traversal methods.
This package was used as a lightweight substitute for NetworkX when building a chain of reasoning using the pseudo-MPE algorithm (see \ref{subsubsec:pseudo-mpe} ).
As the \texttt{MPETree} object is guaranteed to be a single chain, using a simpler resource than NetworkX was advantageous.

@@

The full procedure can be found in \cite{koller2007dseparation}; my implementation shown in \ref{alg:d-separation} follows this pseudocode very closely but the procedure \texttt{d-separated}, instead of finding all nodes $R$ that are d-connected to the input $X$, tests if a given target $Y$ is d-separated from $X$ or not.
This gives some extra flexibility in how the function can be used.
To find the set $S$ of all nodes d-separated from $X$ I simply iterate the \texttt{d-separated} test over all nodes $V$ in the graph representing the BN.

\begin{algorithm}[htp!]
	\caption{d-separation algorithm}
	\label{alg:d-separation}
	\begin{algorithmic}
		\STATE $separated\_list = \emptyset$
		\STATE {\bf input:} source $X$, evidence $E$, nodes $V$
		\FOR{target $Y \in V$} 
		\STATE append $d-separated(X, Y, E)$ to $separated\_list$ \COMMENT{will return true or false}
		\ENDFOR
		\STATE {\bf output:} $separated\_list$
	\end{algorithmic}
\end{algorithm}

### 2019-07-03 methodology/novel-contributions.tex
.
Integrati TODO di AA per Cap. 3 e sezione su correlazione e IG

\subsection{Algorithms}
An important part of my work was developing the algorithms needed to adapt the ideas presented in the paper \enquote{Explaining the Most Probable Explanation} by \cite{Butz2018} and \enquote{A Progressive Explanation of Inference in \enquote{Hybrid} Bayesian} by \cite{Kyrimi2016}.
From the former, the construction of the probability tree through a constructive dialogue with the domain expert, the building of counterfactual explanation branches, the automatic generation of the most probable probability tree from initial evidence.
From the latter, the generation of an \enquote{Inverse explanation}.
Finally, an procedure to output a natural language explanation.

\subsubsection{\enquote{Pseudo-MPE}}
\todo{dove critico il fatto che non calcolano veramente l'MPE come dicevano? qui o in cap.4?}

\subsubsection{Alternative explanation branches}

\subsubsection{\enquote{Pseudo-MPE} from random evidence}

\subsubsection{Inverse explanation}
\todo{da fare e trovare nome migliore}

\subsubsection{Natural language explanation}
\todo{magari pensare anche a visual explanations?}

>>>>

\subsection{Algorithms}
An important part of my work was developing the algorithms needed to adapt the ideas presented in the paper \enquote{Explaining the Most Probable Explanation} by \cite{Butz2018} and \enquote{A Progressive Explanation of Inference in \enquote{Hybrid} Bayesian} by \cite{Kyrimi2016}.
From the former, the construction of the probability tree through a constructive dialogue with the domain expert, the building of counterfactual explanation branches, the automatic generation of the most probable probability tree from initial evidence.
From the latter, the generation of an \enquote{Inverse explanation}.
Finally, a simple procedure to output a natural language explanation was developed.

\subsubsection{\enquote{Pseudo-MPE}} \label{subsubsec:pseudo-mpe}
\todo{dove critico il fatto che non calcolano veramente l'MPE come dicevano? qui o in cap.4?}
\todo{eseguibile in modalita' esaustiva, d-separata o thresholdata}

\subsubsection{Alternative Explanation Branches}

\subsubsection{\enquote{Pseudo-MPE} from Random Evidence} \label{subsubsec: pseudo-mpe-random}

\subsubsection{Inverse Explanation}
\todo{da fare e trovare nome migliore}

\subsubsection{Natural Language Explanation}
\todo{da fare e magari pensare anche a visual explanations?}

\subsubsection{Pairwise Correlations}
An interesting addition is an algorithm to measure the interrelatedness between pairs of variables.

\subsubsection{Plot of network} \label{subsubsec:plot-model}
